A Model for Extracting Time Delays from Noisy Time Series Data

### 2013-08-14 doc/Paper1/ms.tex
SOme edits

@@

\section{The Model for the Inference}
We begin by defining our model for the prior knowledge about how the time
delays give rise rise to the observed data. This results in a probability
distribution for the data given some unknown parameters, i.e. the ``sampling
distribution'' (actually distributions, one for each value of the parameters)
$p(D|\theta)$. Once specific data is in hand, the sampling
distributions become a likelihood function for the parameters which can be
combined with a prior distribution $p(\theta)$ to give the posterior distribution.
The posterior distribution for some parameters $\theta$ given data $D$
is given by Bayes' rule:
\begin{equation}
p(\theta|D = D^*) \propto p(\theta)p(D|\theta)|_{D = D^*}
\end{equation}
where $D^*$ is the actual data set that was observed. Throughout this paper,
the data $D$ are a set of measured magnitudes of multiple images of a QSO; see
Figure~\ref{simdata} for an example data set.

The intrinsic quasar variability, in magnitudes, and in whatever band the
observations are carried out in, can be described by a function $q(t)$. For a
two-image system, the resulting light curves, if observed without gaps and
noise, would be described by the functions
\begin{eqnarray}
y_1(t) &=& m_1 + f(t) + \mu_1(t) \\
y_2(t) &=& m_2 + f(t - \tau_2) + \mu_2(t) \\
\end{eqnarray}
where $\{m_1, m_2\}$ are the mean magnitudes of the images,
$\tau_2$ is the time delay of image 2 relative to image 1, and the
functions $\{\mu_1(t), \mu_2(t)\}$ describe any perturbations that are not
shared by both images. A major source of perturbations is the effect of
microlensing on each of the light curves. The data then consist of
noisy, irregularly sampled measurements of the light curves $\{y_1(t), y_2(t)\}$.

@@

The sampling distributions (and hence the likelihood function) will then be
written as a probability distribution for the measurements $\boldsymbol{D}$
given the time delay $\tau$ along with some nuisance parameters.

@@

>>>>

\newcommand{\params}{\boldsymbol{\theta}}
\newcommand{\data}{\boldsymbol{D}}

@@

\section{Bayesian Inference}
For inferring unknown parameters $\params$ from data $\data$,
Bayes' rule gives the posterior distribution
\begin{eqnarray}
p(\params | \data) \propto p(\params)p(\data|\params)
\end{eqnarray}
The prior distribution $p(\params)$ describes prior knowledge or uncertainty
about the values of the parameter(s). The ``sampling distributions''
$p(\data|\params)$
describes prior knowledge about the data we expect to observe, as a function of
$\params$ (Caticha). When $\data$ is fixed $p(\data|\params)$ is a function of
$\params$ only and is called the likelihood function.

We begin by defining our model for the prior knowledge about how the time
delays give rise rise to the observed data. This will determine the sampling
distribution, and also the set of parameters (including nuisance parameters) to
be inferred.

\section{Sampling Distributions}
The intrinsic quasar variability, in magnitudes, and in whatever band the
observations are carried out in, can be described by a function $q(t)$. For a
two-image system, the resulting light curves, if observed without gaps and
noise, would be described by the functions
\begin{eqnarray}
y_1(t) &=& m_1 + f(t) + \mu_1(t) \\
y_2(t) &=& m_2 + f(t - \tau_2) + \mu_2(t) \\
\end{eqnarray}
where $\{m_1, m_2\}$ are the mean magnitudes of the images,
$\tau_2$ is the time delay of image 2 relative to image 1, and the
functions $\{\mu_1(t), \mu_2(t)\}$ describe any perturbations that are not
shared by both images. A major source of perturbations is the effect of
microlensing on each of the light curves. The data then consist of
noisy, irregularly sampled measurements of the light curves $\{y_1(t), y_2(t)\}$.

@@

The sampling distributions (and hence the likelihood function) will then be
written as a probability distribution for the measurements $\data$
given the time delay $\tau_2$ along with some nuisance parameters.

@@

{RANT ABOUT HOW THE GP ASSUMPTION FOR "MICROLENSING" IS A DESCRIPTION OF OUR
PRIOR UNCERTAINTY ABOUT THE SIGNAL **IN OUR PARTICULAR DATA SET**}
